\TESS Photometric Mapping of a Terrestrial Planet in the Habitable Zone: Detection of Clouds, Oceans, and Continents

### 2019-03-25 tex/ms.tex
more text

@@

We downloaded 
\footnote{\url{https://archive.stsci.edu/missions/tess/models/}}
the Solar System and \tess ephemeris data for
Sectors 1 and 2 and used \spiceypy \citep{Acton1996, Acton2017, Annex2017}
to convert the TDB timestamps to 
ephemeris time (ET). We then used the \textsf{spkezr} utility to compute
the relative positions of Sol, Sol d, and \tess for every cadence
in our dataset in the J2000 equatorial reference frame. In this right-handed
coordinate frame, Sol d is centered at the origin, the $y$ axis points along the 
planet's spin axis, and the $x$ axis points toward the vernal equinox. We
do not apply any light travel correction, as the expected delay is on the order
of a few seconds.

@@

\subsection{Systematics Modeling}

While the astrophysical signal that we wish to analyze is present in all \todo{x} 
extracted lightcurves, its strength is modulated by time-variable systematics 
that are correlated but not identical across all sources. 
\todo{The observed lightcurves can be represented as... (systematics x 
Earthshine equation)}

We choose to model this systematic component using Pixel Level Decorrelation 
(PLD). \todo{(details, math, and a figure of the bases (possibly two-panel with 
comparison to telescope housekeeping variables?)}

\subsection{Reflected-Light Mapping with \starry}

\begin{figure}[ht!]
    \begin{centering}
    \includegraphics[width=0.5\linewidth]{figures/geometry.pdf}
    \caption{\label{fig:geometry}
             Geometry of the \starry model. In this frame, the $y$ axis points up, 
             the $x$ axis points to the right, and the $z$ axis points out of the page.
             The planet has unit radius and sits at the origin, while
             the illumination source is along the $y-z$ plane, with the sub-stellar point is
             marked by a dot. The semi-major axis of the terminator is unity, and
             the semi-minor axis of the terminator is denoted $b$. The night side
             is colored black.
             \codelink{Geometry}
             }
    \end{centering}
\end{figure}

@@

\todo{(brief recap of \starry)}

In \cite{Luger2019} we described how to transform between spherical
harmonic coefficients and coefficients in the polynomial basis $\tilde{p}$, whose
terms are of the form $x^i y^j z^k$ for (non-negative) integer $i$ and $j$ 
and $k = 0$ or $1$. Given a representation of a surface map in terms of
a set of coefficients in $\tilde{p}$, we showed how to compute the visible
flux by integrating each of the terms in $\tilde{p}$ across the visible
surface of the projected disk. For unocculted spheres observed in emitted
light, the area of integration was the full disk, which made the integration
trivial: the flux contribution from each term in $\tilde{p}$ was a constant
easily computed via recursion relations involving factorials.

In the present work, we are interested in \emph{reflected} light, so we must
account for the illumination by the star, which varies smoothly
across the dayside of the planet but whose derivative discontinuously changes at the 
terminator, beyond which the illumination is zero everywhere. This requires
two modifications to the \starry algorithm: we must (1) weight all integrands
by the illumination function on the dayside, and (2) modify the integration
limits to truncate the integrals at the day/night terminator.

Let us consider a right-handed coordinate frame in which the planet has unit 
radius and is located at the origin, the illumination source is along the $y-z$
plane, and the sub-stellar point is on the $+y$ axis at $x = 0$
(see Figure~\ref{fig:geometry}). In this frame, the terminator is a segment of
an ellipse, with semi-major axis $a=1$ aligned with the $x$ axis and
semi-minor axis $b$. It is straightforward to show that the sub-stellar point
is located at $y=\sqrt{1 - b^2}$ and the illumination is given by
%
\begin{align}
    I(b; x, y) &= 
    \begin{dcases}
        \sqrt{1 - b^2}y - bz & 
            \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
            y \geq -b \sqrt{1 - x^2}
        \\
        0 & 
            \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
            \mathrm{otherwise}
    \end{dcases}
    \quad ,
    \label{eq:illum}
\end{align}
%
where $z = \sqrt{1 - x^2 - y^2}$.
%
Since the illumination function is just a polynomial in $x$, $y$, and $z$,
weighting the terms in $\tilde{p}$ by this function keeps all terms within the
polynomial basis.
Provided we always rotate the problem such that the
sub-stellar point lies along the $+y$ axis as above, the limits of integration
are $-1 \leq x \leq 1$ and $b\sqrt{1 - x^2} \leq y \leq \sqrt{1 - x^2}$.
Computing the total reflected flux for any surface map is therefore a matter
of solving integrals of the form
%
\begin{align}
    J_{ijk}(b) &= \int_{-1}^{1} \int_{b\sqrt{1 - x^2}}^{\sqrt{1 - x^2}} x^i y^j z^k \mathrm{d} y \ \mathrm{d} x
    \nonumber\\
    &=
    \frac{
        \Gamma\left(\frac{i + 1}{2}\right) \Gamma\left(\frac{j + 1}{2}\right)
    }{
        2 \Gamma\left(\frac{i + j + k + 4}{2}\right)
    }
    \begin{dcases}
        0
        &
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        i \, \mathrm{odd}
        %
        \\
        %
            1 - b^{j + 1}
        & 
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        k = 0
        %
        \\
        %
        \frac{\sqrt{\pi}}{2} - b^{j + 1}\Gamma\left(\frac{j + 4}{2}\right)
        {_2\tilde{F}_1} \left( -\frac{1}{2}, \frac{j + 1}{2}; \frac{j + 3}{2}; b^2\right)
        & 
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        k = 1
    \end{dcases}
    \quad ,
    \label{eq:integral}
\end{align}
%
where ${_2\tilde{F}_1}$ is the regularized hypergeometric function. For all values
of $i$, $j$, and $k$ in $\tilde{p}$, ${_2\tilde{F}_1}$ reduces to simple
trigonometric functions involving $b$. Moreover, it is straightforward to
derive recursion relations for $J_{ijk}(b)$, making its evaluation extremely fast.
The algorithm for computing $J_{ijk}(b)$ is implemented in the development version
of \starry and will be discussed further in upcoming work.

@@

>>>>

\todo{TJD = BJD - 2457000}

@@

We downloaded 
\footnote{\url{https://archive.stsci.edu/missions/tess/models/}}
the Solar System and \tess ephemeris data for
Sectors 1 and 2 and used \spiceypy \citep{Acton1996, Acton2017, Annex2017}
to convert the TJD timestamps to 
ephemeris time (ET). We then used the \textsf{spkezr} utility to compute
the relative positions of Sol, Sol d, and \tess for every cadence
in our dataset in the J2000 equatorial reference frame. In this right-handed
coordinate frame, Sol d is centered at the origin, the $y$ axis points along the 
planet's spin axis, and the $x$ axis points toward the vernal equinox. We
do not apply any light travel correction, as the expected delay is on the order
of a few seconds.

@@

\subsection{Systematics Modeling}
\label{sec:systematics}

While the astrophysical signal that we wish to analyze is present in all 
extracted lightcurves, its strength is modulated by time-variable systematics 
that are correlated but not identical across all postage stamps, since the
illumination of the CCD by Sol d is strongly inhomogeneous. A physical
model of the reflections and scattering that gives rise to the
signal of Sol d on the \tess detector is beyond the scope of this work. Instead,
we treat this as a de-trending problem, where the signal we wish to fit is
shared by all postage stamps, but multiplied by a systematic signal that is
different for each target. This problem is precisely what the technique of
pixel-level decorelation \citep[PLD;][]{Deming2015, Luger2016, Luger2018a}
seeks to address. In PLD, one computes the sum over all light curves at each
point in time, and uses this quantity to normalize each of the individual
light curves. Since (by assumption) each light curve is the product of the astrophysical
signal and the systematics signal, this procedure divides out the astrophysical
signal, producing a basis of vectors that contain \emph{only} functions of the
systematics. This now serves as a ``clean'' basis one can then use to fit the
data, with minimal risk of fitting out astrophysical signals.

We therefore construct a design matrix $\mathbf{B}$ out of the PLD basis. To
increase the flexibility of the systematics model, we append to $\mathrm{B}$ 5 orthogonal
polynomials in time for each of the two orbits of \tess in each sector.
The systematics model $\mathbf{b}$ for the $n^\mathrm{th}$ signal is thus
%
\begin{align}
\mathbf{b}_n = \mathbf{B} \mathbf{w}_n
\end{align}
%
where $\mathbf{w}_n$ are the weights of each regressor.
Since we solve for different weights for each of the 75 light curves in Sector 1
and 20 light curves in Sector 2, our systematics model has
$75 \times (75 + 2 \times 5) + 20 \times (20 + 2 \times 5) = 6,975$ free parameters.
Since the total number of data points is $912,605$, overfitting is not particularly
concerning, but we nevertheless impose an L2 prior on the weights with
standard deviation $\sigma_B = 0.1$.

\subsection{Reflected-Light Mapping with \starry}
\label{sec:starry}

We model the astrophysical component of the signal---the rotational modulation
of Sol d in reflected light---using the \starry code package \citep{Luger2019}.
\starry computes phase curves and occultation light curves for bodies whose
surfaces are expressed as a sum of spherical harmonics. 
The algorithm works by first transforming
from spherical harmonic coefficients to coefficients in the polynomial basis $\tilde{p}$, whose
terms are of the form $x^i y^j z^k$ for (non-negative) integer $i$ and $j$ 
and $k = 0$ or $1$. \citet{Luger2019} showed how to compute the visible
flux by integrating each of the terms in $\tilde{p}$ across the visible
surface of the projected disk. For unocculted spheres observed in emitted
light, the area of integration is the full disk, which makes the integration
trivial: the flux contribution from each term in $\tilde{p}$ is a constant
easily computed via recursion relations involving factorials. The algorithm is
thus extremely fast and very precise.

However, in the present work, we are interested in \emph{reflected} light, so we must
account for the illumination by the star, which varies smoothly
across the dayside of the planet but whose derivative discontinuously changes at the 
terminator, beyond which the illumination is zero everywhere. This requires
two modifications to the \starry algorithm: we must (1) weight all integrands
by the illumination function on the dayside, and (2) modify the integration
limits to truncate the integrals at the day/night terminator.

\begin{figure}[ht!]
    \begin{centering}
    \includegraphics[width=0.4\linewidth]{figures/geometry.pdf}
    \caption{\label{fig:geometry}
             Geometry of the \starry model in reflected light. In this frame, 
             the $y$ axis points up, 
             the $x$ axis points to the right, and the $z$ axis points out of the page.
             The planet has unit radius and sits at the origin, while
             the illumination source is along the $y-z$ plane, with the sub-stellar point is
             marked by a dot. The semi-major axis of the terminator is unity, and
             the semi-minor axis of the terminator is denoted $b$. The night side
             is colored black.
             \codelink{Geometry}
             }
    \end{centering}
\end{figure}

Let us consider a right-handed coordinate frame in which the planet has unit 
radius and is located at the origin, the illumination source is along the $y-z$
plane, and the sub-stellar point is on the $+y$ axis at $x = 0$
(see Figure~\ref{fig:geometry}). In this frame, the terminator is a segment of
an ellipse, with semi-major axis $a=1$ aligned with the $x$ axis and
semi-minor axis $b$. It is straightforward to show that the sub-stellar point
is located at $y=\sqrt{1 - b^2}$ and the illumination is given by
%
\begin{align}
    I(b; x, y) &= 
    \begin{dcases}
        \sqrt{1 - b^2}y - bz & 
            \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
            y \geq -b \sqrt{1 - x^2}
        \\
        0 & 
            \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
            \mathrm{otherwise}
    \end{dcases}
    \quad ,
    \label{eq:illum}
\end{align}
%
where $z = \sqrt{1 - x^2 - y^2}$.
%
Since the illumination function is just a polynomial in $x$, $y$, and $z$,
weighting the terms in $\tilde{p}$ by this function keeps all terms within the
polynomial basis.
Provided we always rotate the problem such that the
sub-stellar point lies along the $+y$ axis as above, the limits of integration
are $-1 \leq x \leq 1$ and $b\sqrt{1 - x^2} \leq y \leq \sqrt{1 - x^2}$.
Computing the total reflected flux for any surface map is therefore a matter
of solving integrals of the form
%
\begin{align}
    J_{ijk}(b) &= \int_{-1}^{1} \int_{b\sqrt{1 - x^2}}^{\sqrt{1 - x^2}} x^i y^j z^k \mathrm{d} y \ \mathrm{d} x
    \nonumber\\
    &=
    \frac{
        \Gamma\left(\frac{i + 1}{2}\right) \Gamma\left(\frac{j + 1}{2}\right)
    }{
        2 \Gamma\left(\frac{i + j + k + 4}{2}\right)
    }
    \begin{dcases}
        0
        &
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        i \, \mathrm{odd}
        %
        \\
        %
            1 - b^{j + 1}
        & 
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        k = 0
        %
        \\
        %
        \frac{\sqrt{\pi}}{2} - b^{j + 1}\Gamma\left(\frac{j + 4}{2}\right)
        {_2\tilde{F}_1} \left( -\frac{1}{2}, \frac{j + 1}{2}; \frac{j + 3}{2}; b^2\right)
        & 
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        k = 1
    \end{dcases}
    \quad ,
    \label{eq:integral}
\end{align}
%
where ${_2\tilde{F}_1}$ is the regularized hypergeometric function. For all values
of $i$, $j$, and $k$ in $\tilde{p}$, ${_2\tilde{F}_1}$ reduces to simple
trigonometric functions involving $b$. Moreover, it is straightforward to
derive recursion relations for $J_{ijk}(b)$, making its evaluation extremely fast.
We implemented the algorithm for computing $J_{ijk}(b)$ in the development version
of \starry
\footnote{\url{https://github.com/rodluger/starry/tree/linear}}
and will be discussed in detail in upcoming work.

@@

As with the systematics model (\S\ref{sec:systematics}), the model
for the astrophysical signal is \emph{linear}: the total signal $\mathbf{a}$ 
is a linear operation on the vector of spherical harmonic coefficients 
that describes the body's surface:
%
\begin{align}
    \mathbf{a} = \mathbf{A} \mathbf{y}
\end{align}
%
where $\mathbf{A}$ is the design matrix \citep[equal to the quantity 
$\mathbf{s}^\top\mathcal{A}\mathbf{R}'\mathbf{R}$ in][]{Luger2019}
and $\mathbf{y}$ is the vector of spherical harmonic coefficients. As we will
see next, the linearity of the problem makes it easily invertible.

@@

Our full model for the flux in the $n^\mathrm{th}$ postage stamp is thus
%
\begin{align}
\mathbf{m}_n = (\mathbf{r}^{-2}) \circ (\mathbf{B} \mathbf{w}_n) \circ (\mathbf{A} \mathbf{y})
\end{align}
%
where $\circ$ denotes the element-wise product of two vectors.
